---
title: "Predict BGC"
author: "Will MacKenzie"
date: "04/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(RGtk2)
require(plyr)
require(rChoiceDialogs)
require(data.table)
require(doBy)
require(utils)
require(labdsv)
require(tools)
require(svDialogs)
require(tcltk)
require(randomForest)
require(foreach)
require(dplyr)
require(reshape2)
require(reshape)
library(doParallel)
require(data.table)
library(MASS)
library(scales)
library(stats)
library(rgl)
library(RColorBrewer)
library(FNN)
library(igraph)
library(raster)
library(maps)
library(mapdata)
library(maptools)
library(sp)
library(colorRamps)
library(rgeos)
library(rgdal)
library(foreign)
require(data.table)
require(ranger)
require(tidyverse)
require(plotKML)

```

# Set analysis Parameters

```{r set analysis parameters}
grid <- "BC2kmGrid"

GCMs <- c("ACCESS1-0", "CanESM2", "CCSM4", "CESM1-CAM5", "CNRM-CM5", "CSIRO-Mk3-6-0", 
          "GFDL-CM3", "GISS-E2R", "HadGEM2-ES", "INM-CM4", "IPSL-CM5A-MR", "MIROC-ESM", 
          "MIROC5", "MPI-ESM-LR", "MRI-CGCM3")

addVars <- function(dat){
  dat$PPT_MJ <- dat$PPT05 + dat$PPT06  # MaY/June precip
  dat$PPT_JAS <- dat$PPT07 + dat$PPT08 + dat$PPT09  # July/Aug/Sept precip
  dat$PPT.dormant <- dat$PPT_at + dat$PPT_wt  # for calculating spring deficit
  dat$CMD.def <- 500 - (dat$PPT.dormant)  # start of growing season deficit original value was 400 but 500 seems better
  dat$CMD.def[dat$CMD.def < 0] <- 0  #negative values set to zero = no deficit
  dat$CMDMax <- dat$CMD07
  dat$CMD.total <- dat$CMD.def + dat$CMD
  X1$CMD.grow <- X1$CMD05 + X1$CMD06 +X1$CMD07 +X1$CMD08 +X1$CMD09
  X1$DD5.grow <- X1$DD5_05 + X1$DD5_06 + X1$DD5_07 + X1$DD5_08 + X1$DD5_09
  X1$CMDMax <- X1$CMD07 # add in so not removed below
  X1$DDgood <- X1$DD5 - X1$DD18
  X1$DDnew <- (X1$DD5_05 + X1$DD5_06 +X1$DD5_07  + X1$DD5_08)  - (X1$DD18_05 + X1$DD18_06 +X1$DD18_07 +X1$DD18_08)
  X1$TmaxJuly <- X1$Tmax07
  return(dat)
}


### Load random forest model
model <- "16Var_6_2"
fname <- "inputs/models/WNAv11_16_VAR_SubZone_ranger.Rdata"
load(fname)
#rownames(importance(BGCmodel)) ### shows the variable used in the RFmodel


GCMs <- c("ACCESS1-0", "CanESM2", "CCSM4", "CESM1-CAM5", "CNRM-CM5", "CSIRO-Mk3-6-0", 
          "GFDL-CM3", "GISS-E2R", "HadGEM2-ES", "INM-CM4", "IPSL-CM5A-MR", "MIROC-ESM", 
          "MIROC5", "MPI-ESM-LR", "MRI-CGCM3")
rcps <- c("rcp45", "rcp85")
proj.years <- c(2025, 2055, 2085)
hist.years <- c(1995, 2004, 2005, 2009, 2014, 2017)
edatopes <- c("B2", "C4", "D6")
spps.lookup <- fread("inputs/Tree speciesand codes_2.0_2May2019.csv")
edatope.name <- c("Subxeric-poor", "Mesic-medium", "Hygric-rich")
BGCcolors <- fread("inputs/BGCzone_Colorscheme.csv")
```

# ===============================================================================
# generate the vector of mapped BGCs
# ===============================================================================

points <- fread(paste("inputs/", grid, ".csv", sep = ""))
BGC <- points$ID2
table(BGC)

BGC <- gsub(" ", "", BGC)
table(BGC)
sort(table(BGC))

# # merge small units BGC[which(BGC=='CMAwh')] <- 'CMAun'
# BGC[which(BGC=='MHun')] <- 'MHunp' BGC[which(BGC=='ESSFxvw')] <-
# 'ESSFxvp' BGC[which(BGC=='ESSFdcp')] <- 'ESSFdcw'

# BGC zones

zone <- rep(NA, length(BGC))
for (i in BGCcolors$zone){
  zone[grep(i, BGC)] <- i
}
table(zone)


# #===============================================================================
# # select variables from ClimateWNA file and subset by GCM # NB: is
# commented out because you only need to do this once: it just reduces
# ram needed to load in the ClimateBC data
# #===============================================================================
# Columns <- unique(c('PPT05', 'PPT06', 'PPT07', 'PPT08', 'PPT09', 'PPT_at',
# 'PPT_wt', 'CMD07', 'CMD', 'MAT', 'PPT_sm', 'Tmin_wt', 'Tmax_sm',
# rownames(importance(BGCmodel))[-which(rownames(importance(BGCmodel))%in%c('PPT_MJ',
# 'PPT_JAS', 'PPT.dormant', 'CMD.def', 'CMDMax', 'CMD.total'))]))
# 
# #first batch of 8 models fplot=paste('inputs/', grid,
# '_48GCMs_MSYT.csv', sep='') Y1 <- fread(fplot, select = c('GCM',
# Columns), stringsAsFactors = FALSE, data.table = FALSE) #fread is
# faster than read.csv models <-
# c('ACCESS1-0','CanESM2','CCSM4','CESM1-CAM5','CNRM-CM5','CSIRO-Mk3-6-0',
# 'GFDL-CM3','GISS-E2R') for(model in models){ temp <-
# Y1[grep(model,Y1$GCM),] write.csv(temp, paste('inputs/', grid, '_',
# model, '_BioVars.csv', sep=''), row.names = F)
# print(which(models==model)) } #second batch of 7 models
# fplot=paste('inputs/', grid, '_42GCMs_MSYT.csv', sep='') Y1 <-
# fread(fplot, select = c('GCM', Columns), stringsAsFactors = FALSE,
# data.table = FALSE) #fread is faster than read.csv models <-
# c('HadGEM2-ES', 'INM-CM4', 'IPSL-CM5A-MR', 'MIROC-ESM', 'MIROC5',
# 'MPI-ESM-LR','MRI-CGCM3') for(model in models){ temp <-
# Y1[grep(model,Y1$GCM),] write.csv(temp, paste('inputs/', grid, '_',
# model, '_BioVars.csv', sep=''), row.names = F)
# print(which(models==model)) }


# ===============================================================================
# BGC Projections for reference period
# ===============================================================================
vars <- as.data.frame(BGCmodel$variable.importance)
vars <- row.names(vars)
# setwd('C:/GitHub/2019_CCISS')
Columns <- unique(c("PPT05", "PPT06", "PPT07", "PPT08", "PPT09", "PPT_at", 
                    "PPT_wt", "CMD07", "CMD", "MAT", "PPT_sm", "Tmin_wt", "Tmax_sm",
                    vars[!vars %in% c("PPT_MJ", "PPT_JAS", "PPT.dormant", "CMD.def", "CMDMax", "CMD.total")]))

fplot <- paste("inputs/", grid, "_Normal_1961_1990MSY.csv", sep = "")

Y0 <- fread(fplot, select = Columns, stringsAsFactors = FALSE, data.table = FALSE)  #fread is faster than read.csv

Y0 <- Y0[!is.na(Y0[, 2]), ]

Y0 <- addVars(Y0)

Y0 <- Y0 %>% dplyr::select(vars)

## Predict future subzones######
BGC.pred.ref <- predict(BGCmodel, Y0)
#dir.create("./outputs")
write.csv(BGC.pred.ref$predictions, paste("outputs/BGC.pred.ref", grid, "csv", sep = "."), 
          row.names = F)

## Write Climate file ######
fwrite(Y0, paste("inputs/", grid, "_1961_1990_", model,".csv", sep = ""))

# ===============================================================================
# BGC Projections for historical decades
# ===============================================================================

# setwd('C:\\Colin\\Projects\\2019_CCISS')
hist.years <- c(1995, 2005)
hist.periods <- c("1991_2000", "2001_2010")
#hist.year="1995"

for (hist.year in hist.years){
  hist.period <- hist.periods[which(hist.years == hist.year)]
  fplot <- paste("inputs/", grid, "_Decade_", hist.period, "MSY.csv", 
                 sep = "")
  
  Y0 <- fread(fplot, select = Columns, stringsAsFactors = FALSE, data.table = FALSE)  #fread is faster than read.csv
  Y0 <- Y0[!is.na(Y0[, 2]), ]
  # str(Y0)
  Y0 <- addVars(Y0)
  Y0 <- Y0 %>% dplyr::select(vars)
  
  ## Predict future subzones######
BGC.pred <- predict(BGCmodel, Y0)
  
#assign(paste("BGC.pred", hist.year, sep = "."), 
  write.csv(BGC.pred$predictions, paste0("./outputs/BGC.pred", grid, hist.year, ".csv", sep = "."), row.names = F)
  
  ## Write Climate file ######
  write.csv(Y0, paste("inputs/", grid, "_", hist.year, "_", model, ".csv",  sep = ""), row.names = F)
  
  print(hist.year)
}

# ===============================================================================
# BGC Projections for last decade
# ===============================================================================

# setwd('C:\\Colin\\Projects\\2019_CCISS')
fplot <- paste("inputs/", grid, "_2011-2017MSYT.csv", sep = "")

Y0 <- fread(fplot, select = c("ID1", "Year", Columns), stringsAsFactors = FALSE, 
            data.table = FALSE)  #fread is faster than read.csv
# Y0 <- Y0[!is.na(Y0[,2]),]
str(Y0)

Y0 <- addVars(Y0)
Y0 <- Y0 %>% dplyr::select(ID1, Year, vars)
# Extract the year 2017
Y2017 <- Y0[which(Y0$Year == 2017), ]
str(Y2017)

# Calculate mean of 2011-2017 period
Y1 <- Y0 %>%
  group_by(ID1) %>%
  summarise_all(list(mean)) %>%
  ungroup()
##Y1 <- aggregate(Y0, by = list(Y0$ID1), FUN = mean, na.rm = T)[, -1]
Y1 <- Y1[match(Y2017$ID1, Y1$ID1), ]
Y1 <- Y1[,-2]

## Predict BGC units######
BGC.pred.2017 <- predict(BGCmodel, Y2017)
BGC.pred.2014 <- predict(BGCmodel, Y1)
write.csv(BGC.pred.2017$predictions, paste("outputs/BGC.pred", grid, "2017.csv", sep = "."), 
          row.names = F)
write.csv(BGC.pred.2014$predictions, paste("outputs/BGC.pred", grid, "2014.csv", sep = "."), 
          row.names = F)

## Write Climate file ######
fwrite(Y2017, paste("inputs/", grid, "_2017_BioVars.csv", sep = ""))
fwrite(Y1, paste("inputs/", grid, "_2014_BioVars.csv", sep = ""))

# ===============================================================================
# BGC Projections for other historical normals
# ===============================================================================

# setwd('C:\\Colin\\Projects\\2019_CCISS')
hist.years <- c(1995, 2005)
hist.periods <- c("1991_2000", "2001_2010")

mean(c(1991, 2017))
mean(c(2001, 2017))

# read in the data for the 1990s and 2000s
for (hist.year in hist.years){
  hist.period <- hist.periods[which(hist.years == hist.year)]
  fplot <- paste("inputs/", grid, "_Decade_", hist.period, "MSY.csv", 
                 sep = "")
  
  Y0 <- fread(fplot, select = Columns, stringsAsFactors = FALSE, data.table = FALSE)  #fread is faster than read.csv
  Y0 <- Y0[!is.na(Y0[, 2]), ]
  
  Y0 <- addVars(Y0)
  assign(paste("Y", hist.year, sep = "."), Y0)
  print(hist.year)
}

# 2011-2017 period already exported in previous phase, so read that in
Y.2014 <- fread(paste("inputs/", grid, "_2014_BioVars.csv", sep = ""))
Y.2014 <- Y.2014[, -which(names(Y.2014) %in% c("ID1", "Year"))]
str(Y.2014)

Y.2009 <- (Y.2005 * 10 + Y.2014 * 7)/17
Y.2004 <- (Y.1995 * 10 + Y.2005 * 10 + Y.2014 * 7)/27
str(Y.2009)
str(Y.2004)

hist.years <- c(2004, 2009)
for (hist.year in hist.years){
  
  ## Predict future subzones######
  BGC.pred <- predict(BGCmodel, get(paste("Y", hist.year, sep = ".")))
  write.csv(BGC.pred$predictions, paste("outputs/BGC.pred", grid, hist.year, "csv", sep = "."), row.names = F)
  
  ## Write Climate file ######
  write.csv(get(paste("Y", hist.year, sep = ".")), paste("inputs/", grid, 
                                                         "_", hist.year, "_BioVars.csv", sep = ""), row.names = F)
  
  print(hist.year)
}

# ===============================================================================
# BGC Projections for future periods
# ===============================================================================

library(tidyr)
Y0 <- fread(paste0("./inputs/",grid,"_90 GCMsMSY.csv", sep = ""), select = c("Year", "ID1","ID2", Columns))   ##
Y0 <- separate(Y0, Year, into = c("Model","Scenario","FuturePeriod"), sep = "_", remove = T)
Y0$FuturePeriod <- gsub(".gcm","",Y0$FuturePeriod)
Y0 <- Y0 %>% select(Year, ID1,ID2, vars)

require(doParallel)
set.seed(123321)
coreNum <- as.numeric(detectCores()-2)
cl <- makeCluster(coreNum)
registerDoParallel(cl, cores = coreNum)

out <- foreach(GCM = GCMs, .combine = rbind) %:%
  foreach(rcp = rcps, .combine = rbind) %:%
  foreach(proj.year = proj.years, .combine = rbind, .packages = c("data.table","ranger", "dplyr")) %do% {
    sub <- Y0[Y0$Model == GCM & Y0$Scenario == rcp & Y0$FuturePeriod == proj.year,]
    sub <- addVars(sub)
    subPred <- data.frame(ID = sub$ID1)
    subPred$BGC.pred <- predict(BGCmodel, sub)
    
    fwrite(subPred, paste("outputs/BGC.pred", grid, GCM, rcp, proj.year, ".csv", sep = ""))
    temp <- aggregate(ID ~ BGC.pred, data = subPred, FUN = length) %>% 
      mutate(Model = GCM, Scn = rcp, FuturePeriod = proj.year)
    temp
  }



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
